% !TeX root = ../thuthesis-example.tex

\chapter{RESOLUTION OF THESIS DEFENSE COMMITTEE}

Adversarial attack and defense for deep learning is an important and timely topic that concerns the robustness of deep learning in real-world applications. Much work has been done on pattern recognition tasks, while the work on reinforcement learning is less yet increasing. Moreover, there is typically an arm race between attack and defense, thereby calling for comprehensive benchmarks. 

In this thesis, the author systematically studied adversarial robustness on deep reinforcement learning (DRL). It carried out a series of empirical studies of both attack and defense on various video games. Through the experimental results, the author shows the vulnerability of several representative DRL algorithms against attacks that design malicious policies, their transferability, and the ability of image-based attacks to craft effective adversarial observations, even in presence of defense mechanisms.

Through this thesis work, 刘大为 shows his solid knowledge on theoretical fundamentals, his professional knowledge, and his ability to conduct research independently. The thesis is well organized and written. During the oral defense, he presented his work clearly and answered questions correctly.
 
The defense committee voted unanimously in favor of 刘大为’s thesis, and recommended that the degree of Master of science be conferred on him. 


%%This thesis proposes...

%%The language of this section should be in concordance with the original document (either Chinese or English)

%%论文提出了……

%%论文取得的主要创新性成果包括：

%%1. ……

%%2. ……

%%3. ……

%%论文工作表明作者在×××××具有×××××知识，具有××××能力，论文××××，答辩××××。

%%答辩委员会表决，（×票/一致）同意通过论文答辩，并建议授予×××（姓名）×××（门类）学博士/硕士学位。
